---
title: "Task 6"
output:
  html_document:
    df_print: paged
---

```{r, message = FALSE}
library(tidyverse)
library(here)

dog_survey_data <- read_csv(here("data/clean_data/dog_survey_data.csv"))
```

# Dog Survey Response Analysis

## Data Cleaning Process

The dirty data has been passed through a cleaning script which has:

+ removed duplicate rows in the data
+ taken a subset of only the desired columns
+ separated rows where necessary so that each separate dog had its own row
+ standardised the size and gender columns, using patterns to identify and
aggregate common expressions

```{r, echo = TRUE, eval = FALSE}
female_pattern <- "female|girl"

str_detect(dog_gender, female_pattern) ~ "f"
```

+ changed columns into correct type by extracting only the numbers from the
`amount_spent_on_dog_food` column

+ written the clean data to a new file, ready for analysis

## Data Analysis

### Valid email addresses

The client only counts a valid email address as one ending in ‘.com’. How many
survey results have a valid email address?

Creating a column where the value will be TRUE for valid email addresses, and
FALSE for invalid email addresses. valid emails are where they end with ".com"
which is the pattern used in detection. The number of rows can then be counted
and displayed

```{r}
dog_survey_data %>% 
  mutate(valid_email = str_detect(email,".com$")) %>% 
  filter(valid_email == TRUE) %>% 
  nrow()
```

### Average food cost per dog size

What’s the average amount spent on dog food for each dog size?

This can be found by grouping by each dog size and then calculating the average.

```{r}
dog_survey_data %>% 
  group_by(dog_size) %>% 
  summarise(avg_food_cost = mean(amount_spent_on_dog_food, na.rm = TRUE),
            .groups = 'drop_last')
```

### M+ owner's dog ages

For owners whose surname starts with a letter in the second half of the alphabet
(N onwards) what is the average age of their dog?

This can be found by performing a filter and then calculating the average age.

```{r}
dog_survey_data %>% 
  filter(last_name >= "n") %>% 
  summarise(avg_age = mean(dog_age, na.rm = TRUE))

```

### Average age for dogs by gender.

The dog_age column is the age in dog years. The conversion ration is 1 human
year = 6 dog years. Accounting for this, grouping by gender and the average age
for each gender can be calculated.

```{r}
dog_data_accounting_for_age <- dog_survey_data %>% 
  mutate(dog_age = (dog_age/6)) %>%
  group_by(dog_gender) %>% 
  summarise(avg_age = mean(dog_age, na.rm = TRUE), .groups = 'drop_last') %>% 
  drop_na()

dog_data_accounting_for_age
```

### Plot

```{r, eval = TRUE, echo = FALSE}
ggplot()+
  geom_col(data = dog_data_accounting_for_age, aes(x = dog_gender, y = avg_age),
           colour = "black", fill = "grey")

```

