---
title: "Cake Data Analysis"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

```{r, message=FALSE}
library(tidyverse)
library(here)

cake_data <- read_csv(here("data/clean_data/clean_cake_data.csv"))

```

# Cake Data Analysis

The cake data has been cleaned, tidied and combined with the ingredients
information.

### The most cocoa

Which cake has the most cocoa?

cocoa is a possible `ingredient`. Performing a filter for only cakes including
cocoa and then taking a slice of the maximum `liquid_quantity` reveals the cake
with the most cocoa.

```{r}
cake_data %>% 
  filter(ingredient == "cocoa") %>% 
  slice_max(liquid_quantity)
```
###  Cups in a Sponge

How many cups of ingredients go into a sponge cake?

"Sponge cake" exists in the `cake` variable as "Sponge". Sponge cake ingredients
that come in cup measurements can be filtered and then totalled:

```{r}
cups_from_cups <- cake_data %>% 
  filter(cake == "Sponge" & measure == "cup") %>% 
  summarise(sum(quantity)) %>% 
  pull()
```

Alternatively the total number of teaspoons can be calculated, and then divided
by 48 for the total cups from all liquid ingredients in a sponge cake.
```{r}
cups_from_all <- cake_data %>% 
  filter(cake == "Sponge") %>% 
  summarise(sum(liquid_quantity, na.rm = TRUE)) %>% 
  pull()

cups_from_all / 48
```

### Teaspoon ingredients

How many ingredients can be measured in teaspoons?

This can be found by first filtering for "teaspoon" in the `measure` column. And
then counting the distinct rows.

```{r}
cake_data %>% 
  filter(measure == "teaspoon") %>% 
  distinct(ingredient) %>% 
  nrow()
```

It might be useful to see the names of the ingredients:
```{r, echo = FALSE, eval = TRUE}
cake_data %>% 
  filter(measure == "teaspoon") %>% 
  distinct(ingredient) %>% 
  pull()
```

### The most unique cake

Which cake has the most unique ingredients?

This can be found by first grouping by `cake` and then using `n_distinct()` to
count only distinct ingredients in each cake.

```{r}
cake_data %>% 
  group_by(cake) %>% 
  summarise(n_unique_ingredients = sum(n_distinct(ingredient)), .groups = 'drop') %>% 
  slice_max(n_unique_ingredients)
```

### Uncommon ingredients

Which ingredients are used only once?

This can be found by adding a column to count the occurence of each ingredient,
filtering for where the new column `n` is 1 and then pulling the ingredients.

```{r}
cake_data %>% 
  add_count(ingredient) %>% 
  filter(n == 1) %>% 
  select(ingredient) %>% 
  pull()
```


