---
title: "Task 5"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

```{r, message = FALSE}
library(tidyverse)
library(ggplot2)

rw_data <- read_csv(here("data/clean_data/clean_rwa.csv"))
```

# Right-Wing Authoritarianism Survey Response Analysis

The RWA survey data has been clean and average right-wing-authoritarian response
score has been calculated. The score indicates on a scale of 1 - 10 how RWA the
participant's responses were, with 1 being not RWA and 10 being heavily RWA. 

### Average RWA score per gender

Gender is one of the collected variables, so they can be grouped together based
on their response and then the average for each gender can be calculated

```{r}
rw_data %>%
  group_by(gender) %>%
  summarise(average_rwai_per_gender = mean(avg_rwa, na.rm = TRUE),
            .groups = 'drop_last')
```

### Average RWAI for left handed vs. right handed people

Whether a participant was right, left, or both handed was recorded. Participants
can be grouped by their "handedness" and then the average average RWA can be
calculated for each handed group.

```{r}
rw_data %>% 
  group_by(handedness) %>% 
  summarise(average_rwai_per_hand_group = mean(avg_rwa, na.rm = TRUE),
            .groups = 'drop_last') %>% 
  drop_na()
```
### Average RWAI per size of family.

I don't know exactly what this question is asking for. I took it to mean:
what is the average RW score per family size, since literaly translated the
question doesn't make sense - i.e. the average family size for each type of
childhood will be the same as the family size, no?

```{r}
rw_data %>%
  group_by(family_size) %>%
  summarise(average_rwai_per_family_size = mean(avg_rwa, na.rm = TRUE),
            .groups = 'drop_last')
```


### Average test time per education level

Does a persons level of education have an impact on how long it took them to
complete the test?

Education is a measured variable in the data, so it can be easily grouped by and
then the average test time for each can be calculated. Here the median average
was preferred over the mean average. This is because a tiny minority of the
participants took an enormous time to complete the test (maybe they left it and
came back to it).

```{r}
rw_data %>%
  group_by(education_level) %>% 
  summarise(avg_time = median(test_elapse_t, na.rm = TRUE),
            .groups = 'drop_last')
```

```{r}
rw_data %>%
  group_by(education_level) %>% 
  summarise(median(survey_elapse_t, na.rm = TRUE), .groups = 'drop_last')
```

```{r}
avg_time_per_ed_level_tibble <- rw_data %>% 
  group_by(education_level) %>% 
  summarise(avg_test_time = median(test_elapse_t, na.rm = TRUE),
            .groups = 'drop_last') %>%
  drop_na()
```

### Plot results

```{r, echo=FALSE, eval=TRUE}
positions <- c("less than high school", "high school", "university degree", "graduate degree")

ggplot() +
  geom_col(data = avg_time_per_ed_level_tibble, aes(x = education_level, y = avg_test_time), colour = "black", fill = "grey") +
  scale_x_discrete(limits = positions) +
  ylim(0,400) +
  ggtitle("Level of education compared with the average time to complete the test")
```



### Average RWAI per age group

What’s the average RWA score for people in different age bands?

This can be answered first by splitting people into the desired age-bands:

- under 18s,
- 18-25 yr olds,
- 26-40 yr olds,
- 41-60 yr olds
- 61+

and then by grouping based on which age band participants fall in and calculating the average for that age band. 
```{r}
avg_per_age_band <- rw_data %>%
  mutate(age_band = case_when(
    age < 18 ~ "under 18",
    age < 26 ~ "18-25",
    age < 41 ~ "26-40",
    age < 61 ~ "41-60",
    age >= 61 ~ "61+"
  )) %>%
  select(participant_no, age_band, avg_rwa) %>% 
  group_by(age_band) %>% 
  summarise(avg_rwa_for_age_band = mean(avg_rwa, na.rm = TRUE), .groups = 'drop_last')
```

The average right wing authoritarian index score can be plotted for each age
band. Again the score relates to a scale of 1-10, with 1 being not RWA and 10
being RWA.
```{r, echo=FALSE, eval =TRUE}
positions <- c("under 18", "18-25", "26-40", "41-60", "61+")

ggplot() +
  geom_col(data = avg_per_age_band,
           aes(x = age_band, y = avg_rwa_for_age_band),
           colour = "black", fill = "grey") +
  scale_x_discrete(limits = positions) +
  ylim(0,10) +
  ggtitle("Right wing score vs age")
```

```{r}
avg_rwa_per_country <- rw_data %>% 
  group_by(country_ip) %>% 
  summarise(average_rwa = mean(avg_rwa, na.rm = TRUE),
            .groups = 'drop_last')
```

```{r, eval=TRUE, echo=FALSE}
avg_rwa_per_country_group1 <- avg_rwa_per_country %>%
  filter(country_ip < "FA")

avg_rwa_per_country_group2 <- avg_rwa_per_country %>% 
  filter(country_ip >= "FA" & country_ip < "LA")

avg_rwa_per_country_group3 <- avg_rwa_per_country %>% 
  filter(country_ip >= "LA" & country_ip < "SA")

avg_rwa_per_country_group4 <- avg_rwa_per_country %>% 
  filter(country_ip >= "SA")
```

```{r, eval=TRUE, echo=FALSE}

ggplot() +
  geom_col(data = avg_rwa_per_country_group4,
           aes(x = country_ip, y = average_rwa),
           colour = "black", fill = "grey") +
  ylim(0,10) +
  ggtitle("Right wing score vs country") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  coord_flip()
```

```{r, eval=TRUE, echo=FALSE, warning=FALSE}

ggplot() +
  geom_col(data = avg_rwa_per_country_group3,
           aes(x = country_ip, y = average_rwa),
           colour = "black", fill = "grey") +
  ylim(0,10) +
  ggtitle("Right wing score vs country") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  coord_flip()
```

```{r, eval=TRUE, echo=FALSE}

ggplot() +
  geom_col(data = avg_rwa_per_country_group2,
           aes(x = country_ip, y = average_rwa),
           colour = "black", fill = "grey") +
  ylim(0,10) +
  ggtitle("Right wing score vs country") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  coord_flip()
```
```{r, eval=TRUE, echo=FALSE}
ggplot() +
  geom_col(data = avg_rwa_per_country_group1,
           aes(x = country_ip, y = average_rwa),
           colour = "black", fill = "grey") +
  ylim(0,10) +
  ggtitle("Right wing score vs country") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  coord_flip()
```

### Average RWA per urban classification

Does the urban classification have an impace on the participants RWAI?
Urban classifications:

- rural,
- urban,
- suburban,

were recorded. The participants can be grouped based on their response and then
the average for each classification can be calculated.
```{r}
rw_data %>% 
  group_by(urban_classification) %>% 
  summarise(mean(avg_rwa, na.rm = TRUE), .groups = "drop_last")
```

