---
title: "Task 1"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

load in the required packages
```{r, message = FALSE}
library(tidyverse)
library(here)
```

read in the clean data
```{r, message = FALSE}
decathalon_data <- read_csv(here("data/clean_data/decathalon_results_data.csv"))
```

# Decathalon Data Analysis

### The longest long jump

`long_jump` is the variable relating to long jumps. Selecting only that variable
and finding the max value:

```{r}
decathalon_data %>% 
  select(name, event, distance_or_time) %>% 
  filter(event == "long_jump") %>% 
  slice_max(distance_or_time, n = 1)
```

### Average 100m time in each competition

There are two competitions in the data. `x100m` is the variable relating to 

```{r}
decathalon_data %>%
  filter(event == "x100m") %>% 
  group_by(competition) %>% 
  summarise(average_100m = mean(distance_or_time, na.rm = TRUE),
            .groups = 'drop_last')
```

### Top total points across both competitions

Who had the highest combined total points from both competitions?

Looking at a subset of the data and then collapsing for unique rows, the
result is one row per athlete per competition. 

```{r}
decathalon_data %>%
  select(name, points, competition) %>%
  unique() %>% 
  group_by(name) %>% 
  summarise(total_points = sum(points), .groups = 'drop') %>% 
  slice_max(total_points) %>% 
  pull(name, total_points)
```

### Top three shot-putters in each competition

Filtering for just the shot-put event and then grouping the athletes into the
two competitions, a slice of the top three in each competition can be taken.
The distance is included.

```{r}
decathalon_data %>%
  filter(event == "shot_put") %>% 
  group_by(competition) %>% 
  slice_max(distance_or_time, n = 3)
```

### 400m time impact on points

How does the average points change based on how the athlete performed in th
400m event. Faster runners are classed as athletes with a time of under 50s, and
slower runners are athletes with a time of over 50s.

```{r}
faster_runners <- decathalon_data %>%
  filter(event == "x400m" & distance_or_time < 50)

slower_runners <- decathalon_data %>%
  filter(event == "x400m" & distance_or_time > 50)

```

The average points for the faster runners:
```{r, echo=TRUE, eval=TRUE}
faster_runners %>% 
  summarise(average_points = mean(points, na.rm = TRUE)) %>% 
  pull()
```

The average points for the slower runners:
```{r, echo=FALSE, eval=TRUE}
slower_runners %>% 
  summarise(average_points = mean(points, na.rm = TRUE)) %>% 
  pull()
```


