---
title: "Task 3"
output:
  html_document:
    df_print: paged
    toc: TRUE
    toc_float: TRUE
---

```{r, message = FALSE}
# load in the required packages
library(tidyverse)
library(here)
library(maps)

# read in the clean data
seabird_data <- read_csv(here("data/clean_data/birds_and_ship_data.csv"))
```

# Seabird Sightings Data Analysis

## Cleaning process

The data for this analysis comes from two sheets in an excel file relating to
seabird sightings and the ships they were sighted from.

### Cleaning the seabirds data:

The seabirds data has been passed through a cleaning script which has:

+ taken a subset of only the data required for analysis: `record`, `record_id`,
`species`, `species_scientific` - which refers to the scientific name for each
spieces -, `species_abbreviation`, and `count`. Only these variables were
retained from the raw seabirds data.

+ brought the column (variable) names listed above into accordance with the
style guide - snake_case, 

+ removed extraneous data relating to
+ + plumage
+ + age
+ + plumage patterns

  from the `species`, `species_scientific`, and `species_abbreviation`
variables with an `str_detect` function, which filtered out the patterns that
were not relevant for the analysis i.e. the distinction between the ages of the
same species of bird is not something that this analysis is concerned with.

+ saved the clean seabirds data to a new file to be used later.

### Cleaning the ships data:

The ships data has been passed through a cleaning script which has:

+ taken only a subset of the initial data required for this analysis. Only:
`record`, `record_id`, and the latitude/longitude columns were needed.

+ brought the column names into accordance with the style guide.

+ saved the clean data to a new file to be used later.

### Combining the birds and ships data:

The clean data for the seabirds sightings was combined with the clean data,
using a join, for the ships they were sighted from resulting in a dataset with
both latitude, longitude, species information, and count which was ready for
analysis. 

### Assumptions:

1. where count was recorded as 0. This has been changed to NA for analysis. It
was decided 0 cannot be counted as 1 or greater than 1.

2. the same classification of species is used throughout. i.e. where AD is
suffixed to indicate the age of a bird. In all cases, any suffix to the species
data which could be considered as a suffix has been removed. (e.g. AD = adult,
PL1 = plumage pattern 1, DRK = dark plumage colour).

+ Reasoning: the distinction of bird age, plumage type, or colour is not required
for the analysis. Including unnecessary data would directly influence the
results.

3. It is assumed that `record`, `record_id`, `count` are all integers

4. It is assumed the data has been recorded correctly.

### Assertions:

+ the columns must be the correct type before analysis,
+ there must be the correct number of columns in each dataset before analysis,
+ lat/longs must be within a certain range - (-90 - 90/ -180 - 180),
+ count must be within a certain range ( >0 ).

## Data Analysis

The data have been cleaned, tidied and joined into one clean dataset. This hass
been loaded in along with the tidyverse.

### Most individual sightings

Which bird had the most individual sightings?

This question can be answered first by filtering where count is equal to 1 since
the question is concerned with individual sightings, not clusters of more than
1. After grouping by species, the sum of all the sightings of each species can
be calculated, put in descending order and then cut after the first row.

```{r}
seabird_data %>% 
  # don't need lat/long cols
  select(-lat, -long) %>% 
  filter(count == 1) %>% 
  group_by(species, species_scientific, species_abbreviation) %>% 
  summarise(individual_sightings = sum(count, na.rm = TRUE), .groups = 'drop_last') %>% 
  arrange(desc(individual_sightings)) %>% 
  head(1)
```

### The most seen

Which bird has the highest total sightings?

After grouping by species, the total count can be calculated in the same way as
above.

```{r}
seabird_data %>% 
  select(-lat, -long) %>%
  group_by(species, species_scientific, species_abbreviation) %>% 
  summarise(sightings = sum(count, na.rm = TRUE), .groups = 'drop_last') %>% 
  arrange(desc(sightings)) %>% 
  head(1)
```

### Most seen above -30 lat 

To find the seabird with the most sightings above -30 lattitude, the process is
above but with a filtration for lat greater than -30.

```{r}
seabird_data %>% 
  select(-long) %>% 
  filter(lat > -30) %>%
  group_by(species, species_scientific, species_abbreviation) %>% 
  summarise(sightings = sum(count, na.rm = TRUE), .groups = 'drop_last') %>% 
  arrange(desc(sightings)) %>% 
  head(1)
  
```


### Number of lone-seabirds

How many different types of birds were only ever seen in groups of 1?

The average no in a sighting for each species can be calculated. After filtering
this new column to where the average no in a sighting is one, the number of rows
can be counted.

```{r}
seabird_data %>% 
  select(-lat, -long) %>%
  group_by(species, species_scientific, species_abbreviation) %>% 
  summarise(avg_sightings = mean(count, na.rm = TRUE), .groups = 'drop_last') %>% 
  filter(avg_sightings == 1) %>% 
  nrow()
```
It might be useful to see which birds they were:
```{r}
seabird_data %>% 
  select(-lat, -long) %>%
  group_by(species, species_scientific, species_abbreviation) %>% 
  summarise(avg_sightings = mean(count, na.rm = TRUE), .groups = 'drop_last') %>% 
  filter(avg_sightings == 1) %>% 
  select(-avg_sightings)
```

### Penguins

How many penguins were seen?

After converting the species column to lower case, the rows containing penguins
can be found.

```{r}
penguin_pattern <- "penguin"

penguins <- seabird_data %>% 
  mutate(species = str_to_lower(species)) %>% 
  filter(str_detect(species, penguin_pattern))
```

```{r}
penguin_sightings <- penguins %>% 
  summarise(sightings = sum(count, na.rm = TRUE), .groups = 'drop_last') %>% 
  arrange(desc(sightings))
```


### Plotting the penguins

First load in the shape of the world map polygon from the maps package.
```{r}
earth_map <- map_data("world")
```

Then plot the polygon and the plot each sighting by its lat/long co-ordinate.

After viewing the whole map, the decision was made to zoom in to where the
sightings were entirely concentrated: the south pacific.
```{r, warning=FALSE}
ggplot() +
  geom_polygon(data = earth_map, aes(x = long, y = lat, group = group),
               fill = "grey", alpha = 1.5) +
  geom_point(data = penguins ,
             aes(x = long, y = lat, colour = species, size = count)) +
  labs(color = "Species",
       size = "Count",
       title = "Penguin Sightings") +
  xlim(75, 175) +
  ylim(-70, 0)
```




