---
title: "Cake Data Analysis"
output:
  html_document:
    df_print: paged
---

```{r, message=FALSE}
library(tidyverse)

cake_data <- read_csv("data/clean_data/clean_cake_data.csv")

```
### Which cake has the most cocoa in it?

```{r}
cake_data %>% 
  filter(ingredient == "cocoa") %>% 
  slice_max(liquid_quantity)
```
###  For sponge cake, how many cups of ingredients are used in total?

```{r}
cups_from_cups <- cake_data %>% 
  filter(cake == "Sponge" & measure == "cup") %>% 
  summarise(sum(quantity)) %>% 
  pull()
```
Alternatively the number of teaspoons can be calculated, and then divided by 48
for the total cups from all ingredients.
```{r}
cups_from_all <- cake_data %>% 
  filter(cake == "Sponge") %>% 
  summarise(sum(liquid_quantity, na.rm = TRUE)) %>% 
  pull()

cups_from_all / 48
```

### How many ingredients are measured in teaspoons?

```{r}
cake_data %>% 
  filter(measure == "teaspoon") %>% 
  distinct(ingredient) %>% 
  nrow()
```
It might be useful to see the names of the ingredients:
```{r, echo = FALSE, eval = TRUE}
cake_data %>% 
  filter(measure == "teaspoon") %>% 
  distinct(ingredient) %>% 
  pull()
```

### Which cake has the most unique ingredients?

```{r}
cake_data %>% 
  group_by(cake) %>% 
  summarise(n_unique_ingredients = sum(n_distinct(ingredient)), .groups = 'drop') %>% 
  slice_max(n_unique_ingredients)
```

### Which ingredients are used only once?

```{r}
cake_data %>% 
  add_count(ingredient) %>% 
  filter(n == 1) %>% 
  select(ingredient) %>% 
  pull()
```


